---
title: "specLTest"
author: "Witold Wolski"
date: "29 Jul 2015"
output: pdf_document
---

```{r insallLatestversions}
#library(devtools)
#install_github("wolski/prozor")
#install_github("fgcz/specL")
rm(list=ls())
library(prozor)
library(specL)
```

```{r prepareEnvLoadLibraries, echo=FALSE}
R.Version()
library(specL)
packageVersion('specL')
```

Define the parameter of the library generation

```{r setparameters}
if(!exists("OUTPUTDIR")){
  OUTPUTDIR = "res.22348"
  WORKDIR = "temp.22348"
  FASTA_FILE <- "temp.22348/fgcz_9606_d_20140506.fasta"
  MAX_IONS <- 6
  MIN_IONS <- 6
  BLIB_FILTERED <- "blib.blib.filtered.db"
  BLIB_REDUNDANT <- "blib.blib.db"
  MZ_ERROR <- 0.8 # e.g for Q-Exactive
} else{
  print("NOW I am starting to rock!")
}

SPECLIBRARYRDATA = file.path(OUTPUTDIR, "specLLibrary.RData")
SWATH_LIBRARY <- file.path(OUTPUTDIR, "assay_library.tsv")
PEPPROTMAPPING <- file.path(OUTPUTDIR, "pepprot.tsv")

FRAGMENTIONMZRANGE <- c(300,1250)
FRAGMENTIONRANGE <- c(MIN_IONS,200)

ANNOTATEDRDATA <- file.path(WORKDIR,"Annotated.RData")
NON_REDUNDANT <- file.path(WORKDIR,BLIB_FILTERED)
REDUNDANT <- file.path(WORKDIR,BLIB_REDUNDANT)

cat("Set params SPECLIBRARYRDATA = ",SPECLIBRARYRDATA , "\n",
    " SWATH_LIBRARY = ", SWATH_LIBRARY, "\n", 
    " PEPPROTMAPPING = ",PEPPROTMAPPING , "\n",
    " MZ_ERROR = ",MZ_ERROR, "\n",
    " FRAGMENTIONMZRANGE = ", FRAGMENTIONMZRANGE, "\n",
    " FRAGMENTIONRANGE = ",FRAGMENTIONRANGE, "\n",
    " ANNOTATEDRDATA = ",ANNOTATEDRDATA, "\n",
    " NON_REDUNDANT = ",NON_REDUNDANT , "\n",
    " REDUNDANT = ",REDUNDANT, "\n" ,  
    " FASTA_FILE = ", FASTA_FILE, "\n")

```

## Loading the libraries

```{r readDatabases, eval=TRUE, echo=FALSE}
system.time( nonRedundantBlib <- read.bibliospec(NON_REDUNDANT) )
system.time( redundantBlib <- read.bibliospec(REDUNDANT) )
save(redundantBlib, file=file.path(WORKDIR,"redundant.Rdata"))
save(nonRedundantBlib, file=file.path(WORKDIR,"nonredundant.Rdata"))

```

Number of loaded unique precurosrs is `r length(nonRedundantBlib)`.

## Generate Ion Libraries

Define filtering function.

```{r defineFragmentFunction}
fragmentIonFunctionUpTo3 <- function (b, y) {
  Hydrogen <- 1.007825
  Oxygen <- 15.994915
  Nitrogen <- 14.003074
  b1_ <- (b )
  y1_ <- (y )
  b2_ <- (b + Hydrogen) / 2
  y2_ <- (y + Hydrogen) / 2 
  return( cbind(b1_, y1_, b2_, y2_) )
}

```

```{r generateLibrary,echo=FALSE}
load(file.path(WORKDIR,"redundant.Rdata"))
load(file.path(WORKDIR,"nonredundant.Rdata"))

length(nonRedundantBlib)
length(redundantBlib)


specLibrary <- genSwathIonLib(data = nonRedundantBlib,
  data.fit = redundantBlib,
  max.mZ.Da.error = MZ_ERROR,
  topN = MAX_IONS,
  fragmentIonMzRange = FRAGMENTIONMZRANGE,
  fragmentIonRange = FRAGMENTIONRANGE,
  fragmentIonFUN = fragmentIonFunctionUpTo3)
save(specLibrary , file = SPECLIBRARYRDATA )


```

The size of the generated ion library is :`r length(specLibrary@ionlibrary)`.

```{r printSummary,echo=TRUE}
MySum <- summary(specLibrary)

```

## Assign peptide spectrum matches to proteins using prozor

```{r getPeptideProt}
protpep = getProteinPeptideTable(specLibrary)

```

```{r prozor, echo=FALSE, message=FALSE }

library(prozor)
fasta = read.fasta(file = FASTA_FILE, as.string = TRUE, seqtype="AA")
spid <-grep("sp\\|",names(fasta))
fasta = fasta[spid]
length(fasta)
protpepAnnot = annotatePeptides(protpep,fasta)

write.table(protpepAnnot,file=PEPPROTMAPPING,quote = FALSE, row.names = FALSE)

pepProtMatrix = prepareMatrix(protpepAnnot)
dim(pepProtMatrix)
protPepAssingments = greedy(pepProtMatrix)
protPepAssingments[1:10]

for(i in 1:length(specLibrary@ionlibrary)){
  specl <- specLibrary@ionlibrary[[i]]
  id <- paste(specl@peptideModSeq, specl@prec_z, sep="." )
  tmp = protPepAssingments[[id]]
  if(!is.null(tmp)){
    specLibrary@ionlibrary[[i]]@proteinInformation = tmp
  }
}

```

The `r dim(pepProtMatrix)[1]` peptide precursors matched `r dim(pepProtMatrix)[2]` protein sequences.
The minimal protein set explaining these precursors has a size of `r length(unique(unlist(protPepAssingments)))` proteins.


```{r writeSpecnaut}
write.spectronaut(specLibrary,file=SWATH_LIBRARY)

```
